---
title: "Getting to the Top of Spotify: Analyzing Popularity Trends in Music"
author: "Carol Lei, Dylan Hartono, Whitney Zhang, Dainese Chandra"
date: "5/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("knitr")
library("dplyr")
library("lubridate")
library("leaflet")
library("ggplot2")
source("scripts/summary.R")
source("scripts/summary-table.R")
data <- read.csv("data/top10s.csv", stringsAsFactors = FALSE)
```


## Introduction to the problem domain and dataset
```{r echo = FALSE}
summary <- get_summary_info(data)
```
We love listening to all types of music. Therefore, when we are going to top hits playlists, we wondered why certain songs tended to dominant in the Spotify playlists, so we want to ask the question, "what factors allow songs to be popular?" 
We took to a dataset that was pulled from the Spotify API. The dataset contains `r summary$number_of_songs` songs. For each song, there is information about `r summary$features`. We will use these features to determine a certain feature that impacts popularity directly.

Looking at the data overall, since we are looking at the most popular songs, here is the popularity score to better understand what it means -- `r summary$avg_pop`. There are a lot of genres in this dataset; there are `r summary$genres` genres! The most popular genre, which is determined by the highest average popularity is `r summary$top_genre`. The top artist overall though is determined through the artist with the most songs in the top of Spotify. The top artist is `r summary$top_artist`.


## Summary about Spotify Genres
To get a better idea of popularity specific to genre on Spotify, we aggregated our data and found the number of songs with that genre and the mean of the popularity for each of the genres. BPM and Danceability are also included since they will be used later in visualizations. Count of the songs were included since this is a top songs compilation of data, so more songs in the top ranks of popularity were included more. Here is just a small set of our data since there are many genres.

```{r, echo=FALSE}
kable(head(get_summary_table(data), 15),
      col.names = c("Top Genre",
                    "Average Popularity",
                    "Total No. Tracks",
                    "Average Danceability",
                    "Average BPM"))
```

Observations:

- Looking at the data, Dance pop has the most number of songs in the Spotify playlists, but it's popularity for each of the songs tend to just be average.

- Most songs actually only have a couple songs in each genre, so the averages might not truly reflect the overall popularity of the genre itself. For instance, brostep has 2 songs with 86 popularity.

- The lowest average popularity song seems to come from Canadian Latin with only one song. The song is 18 popularity points.

- Most songs however seem to land around 60-80 across varying genres.

## Chart 1: 
### This plot was intended to show the correlation between a song's tempo and its popularity.
```{r, echo=FALSE, message=FALSE}
source("scripts/first-chart.R")
first_chart(data)
```
**Note:** Popularity index was taken from spotify's popularity index.The higher the value, the more popular the song was (100 meaning the highest possible level of popularity attainable on Spotify). The grey area indicates all the values (songs with respective bpm and popularity level), with the blue line being the best fit line.


**Observation**:

From the line plot, we can see that as as the popularity level of a song increases, the bpm (tempo) increases, peaking at about 121bpm. 

This relationship changes once we pass a popularity level of 72, where the downward-sloping part of the plot shows that the higher the popularity level, the lower the bpm.
We can see that songs with high popularity level approaching the 100 level have decreasing tempo that approaches 103 bpm.
This implies that songs that are more popular amongst
the general music audience have lower bpm, possiblying showing
that slow songs have been more popular.

We also note that there is a large error margin (seen in the grey area), where
songs of the same popularity can have either very high bpm or very low bpm. As
such, it is important to note that there are many variables which can contribute
a song's popularity. Although there is a correlation between the tempo of a song and 
its popularity, causation is not proven.

## Chart 2: 
We wanted to visualize if popularity had a correlational relationship with the danceability of a track. This scatter plot depicts track popularity compared to track danceability, both measured on a scale from 0 (not popular) to 100 (popular). Each dot on the plot represents a single track by an artist. The dots are colored to represent the various genres of Spotify's top tracks between 2010 and 2019. 

```{r echo = FALSE}
source("scripts/second-chart.R")
data <- read.csv("data/top10s.csv", stringsAsFactors = FALSE)
second_chart(data)
```

**Insights**:

This scatter plot displays a weak to no positive correlation between popularity and danceability of a track. Though there may not be a clear positive correlation, there is most definitely no negative correlation between the two variables. As danceability increases, popularity trends increases. 


## Chart 3: Pop Music Genre Changes from 2010 to 2019
This bar chart was intended to show the most popular pop songs and music genres changes within a decade. Each of on the chart represent the most popular song(s) of that year with the highest popularity index.
```{r, echo=FALSE, message=FALSE}
source("scripts/third-chart.R")
third_chart(data)
```


**Note:** The popularity index was based on spotify's popularity index. Ranging from 0 to 100, which shows the least popular songs to the most popular songs.

**Observation:**

Throughtout years, people's favorite music genres have changed over time. From this horizontal bar chart, we could see that there are four main pop music genres 
that are most popular from 2010 to 2019, which are *canadian contemporary R&B, dance pop, neo mellow, and pop music*. 

From year 2011 to 2013, the most popular pop song genre is the traditional pop music, which has the highest popularity index three years in a row. Notably, the canadian contemporary R&B became the most popular music genre in 2016. In addition, the traditional pop music won the most popular pop music genre of the year 2019, which had the highest popularity index among other music genres on the chart.

