---
title: "index"
author: "team!"
date: "5/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("knitr")
library("dplyr")
library("lubridate")
library("leaflet")
library("ggplot2")
source("scripts/summary.R")
source("scripts/summary-table.R")
data <- read.csv("data/top10s.csv", stringsAsFactors = FALSE)
```


## Introduction to the problem domain and dataset
```{r echo = FALSE}
get_summary_info(data)
kable(get_summary_table(data))
```


## Summary Table 



## Chart 1: 
### This plot was intended to show the correlation between a song's tempo and its popularity.
```{r, echo=FALSE, message=FALSE}
source("scripts/first-chart.R")
first_chart(data)
```
Popularity index was taken from spotify's popularity index.The higher the value, the more popular the song was (100 meaning the highest possible level of popularity attainable on Spotify).


**Observation**:

From the line plot, we can see that as as the popularity level increases,
the bpm decreases, converging at about 115 bpm. This observation changes
when popularity level hits above 50, where there is an upward-sloping
relationship in which the bpm increases with popularity level.

Once we pass a popularity level of 72, we can see that the
higher the popularity level,the lower the bpm.
This can imply that slow songs with a lower bpm have been more popular.

We also note that there is a large error margin (seen in the grey area), where
songs of the same popularity can have either very high bpm or very low bpm. As
such, it is important to note that there are many variables which can contribute
a song's popularity. Although there is acorrelation between the tempo of a song and 
its popularity, causation is not proven.

## Chart 2: 


## Chart 3